\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-03-25 papers/jss/article.Rnw
Ditch the non-portable int64 examples with a description of what the option does.
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@885 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

If we try to set an int64 field in \proglang{R} to double values, we lose
precision:

<<>>=
test <- new(JSSPaper.Example1)
test$repeated_int64 <- c(2^53, 2^53+1)
length(unique(test$repeated_int64))
@

But when the values are specified as character strings, \pkg{RProtoBuf}
will automatically coerce them into a true 64-bit integer types 
before storing them in the Protocol Buffer message:

<<>>=
test$repeated_int64 <- c("9007199254740992", "9007199254740993")
@

When reading the value back into \proglang{R}, numeric types are returned by
default, but when the full precision is required a character value
will be returned if the \code{RProtoBuf.int64AsString} option is set
to \code{TRUE}.  The character values are useful because they can
accurately be used as unique identifiers and can easily be passed to \proglang{R}
packages such as \pkg{int64} \citep{int64} or \pkg{bit64}
\citep{bit64} which represent 64-bit integers in \proglang{R}.

<<>>=
options("RProtoBuf.int64AsString" = FALSE)
test$repeated_int64
length(unique(test$repeated_int64))
options("RProtoBuf.int64AsString" = TRUE)
test$repeated_int64
length(unique(test$repeated_int64))
@

<<echo=FALSE,print=FALSE>>=
options("RProtoBuf.int64AsString" = FALSE)
@

>>>>

On 64-bit platforms, character strings representing large decimal
numbers will be coerced to int64 during assignment to 64-bit protocol
buffer types to work around the lack of native 64-bit types in \proglang{R}.  The
values are stored as distinct int64 values in memory but when accessed
from \proglang{R} language code they will be coerced into numeric values.  If the
full 64-bit precision is required, the \code{RProtoBuf.int64AsString}
option can be set to \code{TRUE} to return int64 values from messages as character
strings.  The character values are useful because they can
accurately be used as unique identifiers and can easily be passed to \proglang{R}
packages such as \pkg{int64} \citep{int64} or \pkg{bit64}
\citep{bit64} which represent 64-bit integers in \proglang{R}.

% test <- new(JSSPaper.Example1)
% test$repeated_int64 <- c(2^53, 2^53+1)
% length(unique(test$repeated_int64))

% test$repeated_int64 <- c("9007199254740992", "9007199254740993")

% options("RProtoBuf.int64AsString" = FALSE)
% test$repeated_int64
% length(unique(test$repeated_int64))
% options("RProtoBuf.int64AsString" = TRUE)
% test$repeated_int64
% length(unique(test$repeated_int64))

% options("RProtoBuf.int64AsString" = FALSE)
